### 2019-08-19 meta_info_collec/metafun/meta2preds.Rmd
meta info vs prediction accuracy fun

---
title: "meta2pred"
author: "Ali"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Meta data vs prediction accuracy

Investigating any relationship between the metadata going into genomes
and the prediction accuracy coming back out

```{r}
source('scripts/import.R')
```

## Reproducibility pred accuracy

comp `predacc4` (trained on Gmax,Creindhardtii,Mpolymorpha,Tcacao)

with `predacc9` (trained on Creinhardtii,Athaliana,Ptrichocarpa,Gmax,Sitalica,Bdistachyon,Slycopersicum,Mpolymorpha,Vvinifera)

```{r}
toplot <- predacc4
colnames(toplot) <- paste(colnames(toplot), '4', sep='.')
tmp <- predacc9
colnames(tmp) <- paste(colnames(tmp), '9', sep='.')
toplot <- merge(toplot, tmp, by.x = 'genome.4', by.y='genome.9')
```

```{r}
library(ggplot2)
bp <- ggplot(toplot, aes(color=factor(trained_on.4), fill=factor(trained_on.9))) +
  scale_fill_manual(values = c("grey55", "red2")) +
  scale_color_manual(values=c("white", "black")) +
  theme_bw() + 
  geom_abline(slope=1, intercept=0, color="grey20")
bp + geom_point(aes(x=acc_overall.4, y=acc_overall.9), shape=21, size=2.5) 
bp + geom_point(aes(x=f1_total_0.4, y=f1_total_0.9), shape=21, size=2.5) 
bp + geom_point(aes(x=f1_total_1.4, y=f1_total_1.9), shape=21, size=2.5) 
```
At a glance, yes the prediction accuracy is reproducible with different training sets,
yes being in the training set is a sort of bonus

## comparisons of high interest

```{r}
metadata$N50frac <- metadata$N50 / metadata$total_len
metadata$busco_C_loss_tran <- metadata$busco_C_tran - metadata$busco_C_geno
metadata$busco_C_loss_prot <- metadata$busco_C_prot - metadata$busco_C_geno

# stat for kmer diversity
k2mers <- metadata[, nchar(colnames(metadata)) == 2]
metadata$k2mer_sd <- apply(k2mers, 1, sd)

prepreg <- merge(metadata, predacc9, by.x='species', by.y='genome')
is_meta <- c(rep(TRUE, ncol(metadata)), rep(FALSE, ncol(predacc9)))

bp <- ggplot(prepreg, aes(color=factor(trained_on))) +
  scale_color_manual(values = c("grey55", "red2")) +
  theme_bw()

# genome size
bp + geom_point(aes(x=total_len, y=acc_overall))
bp + geom_point(aes(x=total_len, y=f1_total_1))
```

```{r}
# N50
bp + geom_point(aes(x=N50, y=acc_overall))
bp + geom_point(aes(x=N50, y=f1_total_1))
```

```{r}
# N50 as a fraction of genome size
bp + geom_point(aes(x=N50frac, y=acc_overall))
bp + geom_point(aes(x=N50frac, y=f1_total_1))
```

```{r}
# GC
bp + geom_point(aes(x=C, y=acc_overall))
bp + geom_point(aes(x=C, y=f1_total_1))
```
```{r}
# busco genomic completeness
bp + geom_point(aes(x=busco_C_geno, y=acc_overall))
bp + geom_point(aes(x=busco_C_geno, y=f1_total_1))
```

```{r}
# busco ratio transcriptome completeness - genome completeness
bp + geom_point(aes(x=busco_C_loss_tran, y=acc_overall))
bp + geom_point(aes(x=busco_C_loss_tran, y=f1_total_1))
```

```{r}
# busco ratio proteom completeness / genome completeness
bp + geom_point(aes(x=busco_C_loss_prot, y=acc_overall))
bp + geom_point(aes(x=busco_C_loss_prot, y=f1_total_1))
```

```{r}
# sd in rel abundance of k2 mers (wondering if we might be measuring,
# say, genome repetitivity / intergenic with the kmers)
bp + geom_point(aes(x=k2mer_sd, y=acc_overall))
bp + geom_point(aes(x=k2mer_sd, y=f1_total_1))
```

## single regression

```{r, results="hide", warning=FALSE}
col_is_na <- apply(apply(prepreg, 1:2, as.numeric), 2, function(x) is.na(x[1]))
to_scale <- (! col_is_na) & is_meta
prepreg_scale <- prepreg
prepreg_scale[, to_scale] <- scale(prepreg_scale[, to_scale])
```

### prep single binomial regression

```{r}
glm_binom <- function(y_frac, x=NULL, weights=prepreg$total_len){
  suc_fail <- round(cbind(y_frac, 1 - y_frac) * weights)
  if (is.null(x)){
    out <- glm(suc_fail ~ 1, family = binomial)
  }else{
    out <- glm(suc_fail ~ x, family = binomial)
  }
  return(out)
}

sigmoid_pred_line <- function(a_model){
  our_fit_fn <- function(x) predict(a_model, newdata = data.frame(x=x), type="response")
  stat_function(fun=our_fit_fn, color="grey30")
}

key_stats <- function(a_model){
  aic <- AIC(a_model)
  coeffs <- coef(a_model)
  p.value <- summary(a_model)$coefficients[2, 4]
  return(c(AIC=aic, coef=coeffs[2], p.value=p.value))
}

```

### vs accuracy
```{r}
xes <- colnames(prepreg_scale[to_scale])
indi_linregs_acc <- lapply(xes, 
                           function(x) glm_binom(y_frac=prepreg_scale$acc_overall, 
                                                 x=prepreg_scale[, x]))
names(indi_linregs_acc) <- xes
intercept_only <- glm_binom(y_frac=prepreg_scale$acc_overall)
indi_linregs_sum <- t(sapply(indi_linregs_acc, key_stats))
# sort by AIC score
indi_linregs_sum <- indi_linregs_sum[order(indi_linregs_sum[,1]), ]
```

#### most significant negative relationships (redundancy reduced)

```{r}
# fragmented buscos in the transcriptome (proteome was second)
bp <- ggplot(prepreg_scale, aes(y=acc_overall, color=factor(trained_on))) +
  scale_color_manual(values=c("black", "red2")) +
  theme_bw()
bp +
  geom_point(aes(x=busco_F_tran)) +
  sigmoid_pred_line(indi_linregs_acc[["busco_F_tran"]])
```

```{r}
# single copy buscos in the transcriptome (proteom was second); AKA lacks alternative splicing
bp + geom_point(aes(x=busco_S_tran)) +
  sigmoid_pred_line(indi_linregs_acc[["busco_S_tran"]])
```

#### most significant positive relationships (redundancy reduced)
```{r}
# duplicated buscos in the transcriptome (proteom was second); AKA has alternative splicing
bp + geom_point(aes(x=busco_D_tran)) +
  sigmoid_pred_line(indi_linregs_acc[["busco_D_tran"]])
```

```{r}
# count of CDS features from gff3 (maybe also related to alternative splicing?)
bp + geom_point(aes(x=CDS)) +
  sigmoid_pred_line(indi_linregs_acc[["CDS"]])
```

```{r}
# count of five_prime_UTR features from gff3 (maybe also related to alternative splicing?)
# that and probably the program that made the gff
bp + geom_point(aes(x=five_prime_UTR)) +
  sigmoid_pred_line(indi_linregs_acc[["five_prime_UTR"]])
# next would be more gff counts: 3', mRNA, exon
```

```{r}
# finally all the total_len_gt_X fall out with x in decreasing order
bp + geom_point(aes(x=total_len_gt_50k)) +
  sigmoid_pred_line(indi_linregs_acc[["total_len_gt_50k"]])
```

## multi regression (PLS after logit function)

```{r}
# linearizing transform for bounded 0-1 data (assuming it followed sigmoid to start with)
logit <- function(x){
  log(x) - log(1-x)
}

prepreg_scale$acc_overall_logit <- logit(prepreg_scale$acc_overall)
```

### OK, once just with lm
```{r}
non_redundant <- c("busco_F_tran","busco_D_tran","CDS","five_prime_UTR","mRNA","total_len_gt_50k","total_len","largest_contig","N75","gene","busco_C_tran","busco_C_loss_tran","N50frac","busco_F_geno","contigs_gt_5k","busco_M_tran","contigs","L50","CC","CA","busco_C_geno","CG","busco_M_geno","C","TA","AA","AC","N","AT","GA","GC","A","busco_D_geno","AG", "k2mer_sd")

x <- prepreg_scale[, non_redundant]
x$y <- prepreg_scale$acc_overall_logit
# bullshit bingo
lm.acc_overall.fit <- lm(y~., data=x)
summary(lm.acc_overall.fit)
```

### Now PLS

```{r}
library(pls)
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on = prepreg_scale$trained_on
x$y <- prepreg_scale$acc_overall_logit
pls.acc_overall.fit <- plsr(y~., data=x)
validationplot(pls.acc_overall.fit)
```

```{r}
# viz prediction accuracy
plot(x$y, predict(pls.acc_overall.fit, x, ncomp=3))
plot(x$y, predict(pls.acc_overall.fit, x, ncomp=12))
```

```{r}
library(reshape2)
metaclass <- c(rep("gff", 6), 
               rep("quast", 18),
               rep("buscoG", 5),
               rep("buscoP", 5),
               rep("buscoT", 5),
               rep("kmers", 13),
               rep("f.eng", 5))
metaclass <- factor(metaclass, levels=unique(metaclass))
# look at loadings of first components
plot_with_loadings <- function(amodel, metaclass){
  pls.loadings <- amodel$loadings
  toplot <- data.frame(pls.loadings[,1:3])
  toplot$meta <- factor(rownames(toplot), levels = rownames(toplot))
  toplot$metaclass <- metaclass
  toplot <- melt(toplot)
  #return(toplot)
  colnames(toplot)[3:4] <- c("component", "loading")

  bp <- ggplot(toplot, aes(y=loading, x=meta, fill=metaclass)) +
    scale_fill_brewer(palette = "Set3") +
    geom_bar(stat="identity") +
    facet_grid(rows=vars(component)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  bp
}
plot_with_loadings(pls.acc_overall.fit, metaclass)
```

#### looking at final coeffs with ncomp=2
```{r}

plot_with_ncomps <- function(ncomps, amodel, metaclass){
  toplot <- coef(amodel, ncomp=ncomps)
  toplot <- data.frame(coeffs3=toplot, 
                       meta=factor(rownames(toplot), 
                                   levels=rownames(toplot)),
                       metaclass=metaclass
                       )
  colnames(toplot)[1] <- 'coeffs'
  ggplot(toplot, aes(x=meta, y=coeffs, fill=metaclass)) +
    geom_bar(stat="identity") +
    scale_fill_brewer(palette = "Set3") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
}
plot_with_ncomps(2, pls.acc_overall.fit, metaclass)
```

#### looking at final coeffs with ncomp=3
```{r}
plot_with_ncomps(3, pls.acc_overall.fit, metaclass = metaclass)
```

#### looking at final coeffs with ncomp=9
```{r}
plot_with_ncomps(9, pls.acc_overall.fit, metaclass)
```

#### looking at final coeffs with ncomp=16
```{r}
plot_with_ncomps(16, pls.acc_overall.fit, metaclass)
```

## OK, and now that we have something resembling useful, PLS f1 1
```{r}
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg_scale$trained_on
x$y <- logit(prepreg_scale$f1_total_1)
pls.f1_total_f1.fit <- plsr(y~., data=x)
validationplot(pls.f1_total_f1.fit)
```

```{r}
# viz prediction accuracy
plot(x$y, predict(pls.f1_total_f1.fit, x, ncomp=3))
plot(x$y, predict(pls.f1_total_f1.fit, x, ncomp=12))
```

```{r}
# loadings
plot_with_loadings(pls.f1_total_f1.fit, metaclass)
```

#### looking at final coeffs with ncomp=2+
```{r}
plot_with_ncomps(2, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(3, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(9, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(16, pls.f1_total_f1.fit, metaclass)
```

## And further trained on 9, PLS f1_total_0
```{r}
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg_scale$trained_on
x$y <- logit(prepreg_scale$f1_total_0)
pls.f1_total_f0.fit <- plsr(y~., data=x)
validationplot(pls.f1_total_f0.fit)
```

```{r}
# viz prediction accuracy
plot(x$y, predict(pls.f1_total_f0.fit, x, ncomp=3))
plot(x$y, predict(pls.f1_total_f0.fit, x, ncomp=12))
```

```{r}
# loadings
plot_with_loadings(pls.f1_total_f0.fit, metaclass)
```

#### looking at final coeffs with ncomp=2+
```{r}
plot_with_ncomps(2, pls.f1_total_f0.fit, metaclass)
plot_with_ncomps(3, pls.f1_total_f0.fit, metaclass)
plot_with_ncomps(9, pls.f1_total_f0.fit, metaclass)
plot_with_ncomps(16, pls.f1_total_f0.fit, metaclass)
```

## kmers...

```{r}
kmers <- prepreg[, nchar(colnames(prepreg)) < 3]
mycols <- c('blue3', 'gold2', 'green3', 'grey40')
boxplot(kmers, col=mycols[c(1,2,4,1,3,3,1,3,2,2,3,2,1)], ylab='rel. freq.', main='kmers')
```

### model without kmers/kmers only (train 9, f1 1)

```{r}
# split input to kmers / no kmers
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg$trained_on
kmerish <- c("gc_content", "A", "C", "AA", "AC", "AG", "AT", "CA", "CC", "CG", "GA", "GC", "TA")
kmers_only <- x[, kmerish]
kmers_only$y <- logit(prepreg_scale$f1_total_1)

no_kmers <- x[, ! colnames(x) %in% kmerish]
no_kmers$y <- logit(prepreg_scale$f1_total_1)
```

#### kmers

```{r}
pls.kmers.f1_total_f1.fit <- plsr(y~., data=kmers_only)
validationplot(pls.kmers.f1_total_f1.fit)
plot(kmers_only$y, predict(pls.kmers.f1_total_f1.fit, kmers_only, ncomp=3))
mck <- metaclass[colnames(x) %in% kmerish]
plot_with_loadings(pls.kmers.f1_total_f1.fit, metaclass=mck)
plot_with_ncomps(2, pls.kmers.f1_total_f1.fit, metaclass=mck)
plot_with_ncomps(3, pls.kmers.f1_total_f1.fit, metaclass=mck)
```

#### no kmers

```{r}
pls.nokmers.f1_total_f1.fit <- plsr(y~., data=no_kmers)
validationplot(pls.nokmers.f1_total_f1.fit)
plot(no_kmers$y, predict(pls.nokmers.f1_total_f1.fit, no_kmers, ncomp=3))

mc2 <- metaclass[! colnames(x) %in% kmerish]
plot_with_loadings(pls.nokmers.f1_total_f1.fit, metaclass=mc2)
plot_with_ncomps(2, pls.nokmers.f1_total_f1.fit, metaclass=mc2)
plot_with_ncomps(3, pls.nokmers.f1_total_f1.fit, metaclass=mc2)
```

## Finally, PLS of f1_total_1 for trained on 4 
(to see if some or all is robust to exact training set)

```{r, results="hide", warning=FALSE}
prepreg <- merge(metadata, predacc4, by.x='species', by.y='genome')
col_is_na <- apply(apply(prepreg, 1:2, as.numeric), 2, function(x) is.na(x[1]))
to_scale <- (! col_is_na) & is_meta
prepreg_scale <- prepreg
prepreg_scale[, to_scale] <- scale(prepreg_scale[, to_scale])
```

```{r}
# these are basically all the input vars
x <- prepreg_scale[, to_scale]
x$trained_on <- prepreg_scale$trained_on
x$y <- logit(prepreg_scale$f1_total_1)
pls4.f1_total_f1.fit <- plsr(y~., data=x)
validationplot(pls4.f1_total_f1.fit)
```
```{r}
# viz prediction accuracy
plot(x$y, predict(pls4.f1_total_f1.fit, x, ncomp=3))
plot(x$y, predict(pls4.f1_total_f1.fit, x, ncomp=12))
```

```{r}
# loadings
plot_with_loadings(pls4.f1_total_f1.fit, metaclass)
```

#### looking at final coeffs with ncomp=2+
```{r}
plot_with_ncomps(2, pls4.f1_total_f1.fit, metaclass)
plot_with_ncomps(3, pls4.f1_total_f1.fit, metaclass)
plot_with_ncomps(9, pls4.f1_total_f1.fit, metaclass)
plot_with_ncomps(16, pls4.f1_total_f1.fit, metaclass)
```

#### and finally side by side, 9 and 4
```{r}
plot_with_ncomps(3, pls.f1_total_f1.fit, metaclass)
plot_with_ncomps(3, pls4.f1_total_f1.fit, metaclass)
```

OK, on the side that just makes sense

* trained_on is positive
* fragmented buscos always hurt
* genomic buscos, sorts: complete, single copy, duplicated missing, fragmented

Makes some sense

* total length hurts

Confusing

* more contigs are good?

Maybe related to calling alternative splicing

* total counts of gff features help, particularly CDS, exon
* duplicated buscos are good, and single copy bad at the transcriptome/proteome level

Worrisome

* kmers

Adding the engineered feature from the kmers (2mer sd) certainly
helps a model without the kmers, but otherwise doesn't capture
that much of the kmer info / replace them.

```{r}
library(ggrepel)
coeffs9 <- coef(pls.f1_total_f1.fit, ncomp=3)
coeffs4 <- coef(pls4.f1_total_f1.fit, ncomp=3)

incl_name <- function(x, keep){
  if (x %in% keep){
    return(x)
  }else{
    return(NA)
  }
}
# prep labels

toorder <- abs(coeffs4 - coeffs9)
toorder <- data.frame(toorder)[order(toorder, decreasing = TRUE),, drop=FALSE]

toplot <- data.frame(coeffs4=coeffs4, coeffs9=coeffs9)
colnames(toplot) <- c("coeffs4", "coeffs9")
toplot$meta <- rownames(toplot)
toplot$meta_labs <- sapply(toplot$meta, incl_name, keep=rownames(toorder)[1:6])

ggplot(toplot, aes(x=coeffs4, y=coeffs9)) + 
  geom_abline(slope=1, intercept=0, color="grey") +
  geom_point() +
  geom_text_repel(aes(label=meta_labs), color="blue") +
  theme_bw()
```